# PRACTICAL MACHINE LEARNING - Jeffrey Leek, JHU
# PREDICTION MODEL Examples

## 1. Email Classification
* Install kernlab package, load spam dataset

```{r}
install.packages("kernlab")
library(kernlab)
data(spam)
head(spam)
```
## Create Simple Frequency Plot 
* Plot occureence of word 'your' in emails

## Simple Prediction Algorithm
•	Find a value C, 
•	If frequency of ‘your’ in email > C, then predict “spam”
*  Add vertical line to demarcate decision boundary

```{r}
plot(density(spam$your[spam$type=='nonspam']), col='blue', main="", xlab="Frequency of 'your'")
lines(density(spam$your[spam$type=="spam"]), col='red')
abline(v=0.5, col='black')
```

## Generate Prediction Table
* Calculate proportion of correct predictions
* Measure of prediction model accuracy

```{r}
prediction <- ifelse(spam$your > 0.5, "spam", "nonspam")
table(prediction, spam$type)/length(spam$type)

```
## Accuracy = 0.459 + 0.292 = 0.751
* The model is 75% accurate
* Optimistic estimate of overall error rate


# Example 2: In-sample versus out of sample error
* Train set versus test set error and model accuracy 

```{r}
library(kernlab); data(spam); set.seed(333); 
smallSpam = spam[sample(dim(spam)[1], size=10),]
spamLabel = (smallSpam$type=="spam")*1 + 1
plot(smallSpam$capitalAve, col=spamLabel)
```
# Email Prediction Models
## Prediction Rule 1: Perfect prediction
* Model reduces in-smaple error rate, 
* but overfits the data

```{r}
rule1 <- function(x){
    prediction <- rep(NA, length(x))
    prediction[x > 2.70] <- "spam"
    prediction[x < 2.40] <- "nonspam"
    prediction[x >= 2.40 & x <= 2.45] <- "spam"
    prediction[x > 2.45 & x <= 2.70] <- "nonspam"
    return(prediction)
}
table(rule1(smallSpam$capitalAve), smallSpam$type)
```
## Prediction Rule 2 - Simple rule
* Higher error rate, but fits the data better
* Misclassifies one 'spam' email as "nonspam"

```{r}
rule2 <- function(x){
    prediction <- rep(NA, length(x))
    prediction[x > 2.40] <- "spam"
    prediction[x < 2.40] <- "nonspam"
    return(prediction)
}
table(rule2(smallSpam$capitalAve), smallSpam$type)
```

## Apply Prediction Model to Complete Spam Data set

```{r}
table(rule1(spam$capitalAve), spam$type)
```


```{r}
table(rule2(spam$capitalAve), spam$type)
```

## Look at Model Accuracy 
* Simple rule is more accurate 

```{r}
sum(rule1(spam$capitalAve)==spam$type)
sum(rule2(spam$capitalAve)==spam$type)
```
## What is overfitting?
* The data has two parts: Signal and noise
* Goal of prediction model is to identify the signal, minimize noise
* Perfect prediction model captures both, but won't predict new data well




