# PRACTICAL MACHINE LEARNING IN R - Jeff Leek
# Week 2.4: Plotting Predictors with Caret package and ggplot2

## EXAMPLE 2: WAGE DATA from ISLR
* Analysis of wage data from Introduction to Statistical Learning
* Install ISLR package, load ggplot2, caret, load wage data 
* NOTE: Sample consists of N=3000 males from Mid-Atlantic region

```{r}
install.packages("ISLR")
library(ISLR); library(ggplot2); library(caret)
data(Wage)
summary(Wage)
```

## Build Training / Test Sets
```{r}
inTrain <- createDataPartition(y=Wage$wage, p=0.7, list=FALSE)

training <- Wage[inTrain, ]
testing <- Wage[-inTrain, ]
dim(training); dim(testing)
```

## Feature plot (caret package)
* All plots constructed using training set
* Select pairplots 

```{r}
#featurePlot(x=training[, c("age", "education", "jobclass")], 
#            y=training$wage, plots="pairs")

pairs(wage ~ age + education + jobclass, data=training);
qplot(age, wage, data=training)
```

## Qplot with color (ggplot2 package)
* Sort jobclass by color
* Separate class of high wage jobs for informatin sector

```{r}
qplot(age, wage, colour=jobclass, data=training)
```

## Add regression smooters (ggplot2 package)
* Sort education by color
* Add geom-smoother for linear model function
* Fits linear model for each education class, with regression line on plot

```{r}
qq = qplot(age, wage, colour=education, data=training)
qq + geom_smooth(method='lm', formula = y~x)
```

## Use cut2 for making factors (Hmisc package)
* Useful to break a variable into different categories, e.g., wage
* load Hmisc, and indicate number of groups 'g' with cut2
* Table shows cutoff points for each category

```{r}
library(Hmisc)
cutWage <- cut2(training$wage, g=3)
table(cutWage)
```
## Create Boxplots with cut2 groupings
* Plot different wage groups in boxplots using qplot
* See relationship between Wage and Age a bit more clearly

```{r}
p1 <- qplot(cutWage, age, data=training, fill=cutWage,
            geom = c("boxplot"))
p1
```

## Boxplots with Points overlayed
* Add actual data points on top of the boxplots
* Large number of dots suggests any trends are real

```{r}
p2 <- qplot(cutWage, age, data=training, fill=cutWage,
            geom = c("boxplot", "jitter"))
grid.arrange(p1,p2, ncol=2)
```

## Tables
* Can also use the factorized version to look at continuous tables of data
* Compare factor version of wages to to jobclass
* More industrial jobs in the lower wage section, more high wage in information sector

```{r}
t1 <- table(cutWage, training$jobclass)
t1
```

###  Use prop.table command to return table of proportions
```{r}
prop.table(t1,1)
```

## Density Plots
* Plot values of continuous predictors as proportion density
* Use qplot() to plot wage distribution by education level
* Good for examining distributions across different groups

```{r}
qplot(wage, colour=education, data=training, geom='density')
```