R Markdown demo Roger R. Peng
=============================

First, load some data

```{r}
library(datasets)
data("airquality")
summary(airquality)
```

# Create pairplots

```{r}
pairs(airquality)
```

# Construct regression model
- Regress Ozone on Wind, Solar Radiation, and Temperature

```{r}
library(stats)
fit = lm(Ozone ~ Wind + Solar.R + Temp, data = airquality)
summary(fit)
```

### Add unordered list using asterisks (lines separated by return)

* First element

* Second element

* etc. 

You can also use LaTeX to add formulas

## Adding Graphics

```{r showtable, results='asis'}
library(xtable)
xt <- xtable(summary(fit))
print(xt, type = 'html')
```

## Setting Global options

```{r setoption, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'hide')
```

First, simulate some data.

```{r simulatedata, echo=TRUE}
x <- rnorm(100)
y <- x + rnorm(100, sd = 0.5)
```

Here is a scatterplot of the data.

```{r scatterplot, fig.height=4}
par(mar = c(5,4,1,1), las=1)
plot(x,y, main = "My Simulated Data")
```

## Common Options: 

### Output
- results: "asis", "hide"
- echo: TRUE, FALSE

### Figures
- fig.height: numeric
- fig.width: numeric


